### 2019-01-19 metrics2.tex
metrics

\begin{enumerate}[(a)]
	\item Denote sample variances as $s^2_{ab}, a, b \in \left\{y, x_1, x_2\right\}$. To compute the OLS estimate we need to solve the following optimization problem:
	\begin{align*}
	\underset{b_0, b_1, b_2}{\min} \summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})^2\\
	\text{FOCs:} \begin{cases}
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i}) = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{1i} = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{2i} = 0
	\end{cases}\\
	b_0 = \bar{y} - b_1\bar{x_1} - b_2 \bar{x_2}\\
	b_1\left(\summa x_{1i}^2\right) = \summa (y_i - \bar{y} + b_1\bar{x_1} + b_2\bar{x_2} - b_2x_{2i})x_{1i}
	\end{align*}
	\begin{align*}
	b_1 = \frac{\summa (y_i - \bar{y} - b_2(x_{2i} - \bar{x}))x_{1i}}{\summa (x_{1i} - \bar{x})^2}&=\frac{\summa (y_i - \bar{y})(x_{1i} - \bar{x_1}) - b_2\summa(x_{2i} - \bar{x_2})(x_{1i} - \bar{x_1})}{\summa (x_{1i} - \bar{x})^2} =\\
	&=\frac{s^2_{yx_1} - b_2s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Similarly for $b_2$:
	\begin{align}\label{eq2}
	b_2 = \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}} &= \frac{s^2_{yx_2} - s^2_{yx_1} + b_2s^2_{x_1x_2}}{s^2_{x_2x_2}} \nonumber\\
	b_2 &= \frac{s^2_{yx_2} - s^2_{yx_1}}{s^2_{x_2x_2} - s^2_{x_1x_2}}
	\end{align}
	\item Performing simple linear regression of $y$ on $x_1$, $y$ on $x_2$, $x_1$ on $x_2$ and $x_2$ on $x_1$ denote OLS estimates for slope coefficients as $\alpha_1, \alpha_2, \alpha_3, \alpha_4$ respectively. It is known that
	\begin{align*}
	\alpha_1 = \frac{s^2_{yx_1}}{s^2_{x_1x_1}},\ \alpha_2 = \frac{s^2_{yx_2}}{s^2_{x_2x_2}},\ \alpha_3 = \frac{s^2_{x_1x_2}}{s^2_{x_2x_2}},\ \alpha_4 = \frac{s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Divide numerator and denumerator of (\ref{eq2}) by $s^2_{x_1x_1}$ (assuming that it is not equal to 0):
	\begin{align*}
	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1}{\frac{\alpha_4}{\alpha_3} - \alpha_4}
	\end{align*}
	Similarly for $b_1$:
	\begin{align*}
	b_1 = \frac{\alpha_1\frac{\alpha_3}{\alpha_4} - \alpha_2}{\frac{\alpha_3}{\alpha_4} - \alpha_3}
	\end{align*}
	\item From (\ref{eq2}):
	\begin{align*}
	\hat{\gamma_2} = \frac{s^2_{yx_2} - s^2_{x_1x_2}}{s^2_{yy} - s^2_{yx_1}} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{yy} - s^2_{yx_1}}{s^2_{yx_2} - s^2_{x_1x_2}}
	\end{align*}
\end{enumerate}
\end{document}

>>>>

\begin{enumerate}[(a)]
	\item Denote sample variances as $s^2_{ab}, a, b \in \left\{y, x_1, x_2\right\}$. To compute the OLS estimate we need to solve the following optimization problem:
	\begin{align*}
	\underset{b_0, b_1, b_2}{\min} \summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})^2\\
	\text{FOCs:} \begin{cases}
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i}) = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{1i} = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{2i} = 0
	\end{cases}\\
	b_0 = \bar{y} - b_1\bar{x_1} - b_2 \bar{x_2}\\
	b_1\left(\summa x_{1i}^2\right) = \summa (y_i - \bar{y} + b_1\bar{x_1} + b_2\bar{x_2} - b_2x_{2i})x_{1i}
	\end{align*}
	\begin{align*}
	b_1 = \frac{\summa (y_i - \bar{y} - b_2(x_{2i} - \bar{x}))x_{1i}}{\summa (x_{1i} - \bar{x})^2}&=\frac{\summa (y_i - \bar{y})(x_{1i} - \bar{x_1}) - b_2\summa(x_{2i} - \bar{x_2})(x_{1i} - \bar{x_1})}{\summa (x_{1i} - \bar{x})^2} =\\
	&=\frac{s^2_{yx_1} - b_2s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Similarly for $b_2$:
	\begin{align}\label{eq2}
	b_2 = \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}} &= \frac{s^2_{yx_2} - s^2_{yx_1} + b_2s^2_{x_1x_2}}{s^2_{x_2x_2}} \nonumber\\
	b_2 &= \frac{s^2_{yx_2} - s^2_{yx_1}}{s^2_{x_2x_2} - s^2_{x_1x_2}}
	\end{align}
	\item Performing simple linear regression of $y$ on $x_1$, $y$ on $x_2$, $x_1$ on $x_2$ and $x_2$ on $x_1$ denote OLS estimates for slope coefficients as $\alpha_1, \alpha_2, \alpha_3, \alpha_4$ respectively. It is known that
	\begin{align*}
	\alpha_1 = \frac{s^2_{yx_1}}{s^2_{x_1x_1}},\ \alpha_2 = \frac{s^2_{yx_2}}{s^2_{x_2x_2}},\ \alpha_3 = \frac{s^2_{x_1x_2}}{s^2_{x_2x_2}},\ \alpha_4 = \frac{s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Divide numerator and denumerator of (\ref{eq2}) by $s^2_{x_1x_1}$ (assuming that it is not equal to 0):
	\begin{align*}
	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1}{\frac{\alpha_4}{\alpha_3} - \alpha_4}
	\end{align*}
	Similarly for $b_1$:
	\begin{align*}
	b_1 = \frac{\alpha_1\frac{\alpha_3}{\alpha_4} - \alpha_2}{\frac{\alpha_3}{\alpha_4} - \alpha_3}
	\end{align*}
	\item From (\ref{eq2}):
	\begin{align*}
	\hat{\gamma_2} = \frac{s^2_{yx_2} - s^2_{x_1x_2}}{s^2_{yy} - s^2_{yx_1}} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{yy} - s^2_{yx_1}}{s^2_{yx_2} - s^2_{x_1x_2}}
	\end{align*}
\end{enumerate}
\section*{Problem 4}
Consider a simple regression model through the origin under the Gauss-Markov assumptions
\begin{align*}
y = \gamma x + \varepsilon
\end{align*}
and let
\begin{align*}
\tilde{\gamma} = \frac{\summa y_ix_i^3}{\summa x_i^4}
\end{align*}
be an estimator of the slope parameter $\gamma$.
\begin{enumerate}[(a)]
\item Under what condition this estimator is well defined?
\item Show conditional unbiasness of $\tilde{\gamma}$.
\item Derive conditional variance of the estimator.
\end{enumerate}

\begin{enumerate}[(a)]
	\item First of all $x_i \neq 0\ \forall i$ then the estimator is well-defined.
	\item \begin{align*}
	\expect[\tilde{\gamma}|x_1, \dots, x_n] = \expect\left[\frac{\summa (\gamma x_i + \varepsilon_i)x_i^3}{\summa x_i^4}\bigg|x_1, \dots, x_n\right] &= \gamma + \frac{1}{\summa x_i^4} \summa x_i^3\expect[\varepsilon_i|x_1, \dots, x_n] =\\
	 &\underset{\text{G.-M. assump}}{=} \gamma + 0
	\end{align*}
	\item \begin{align*}
	Var(\tilde{\gamma}|x_1, \dots, x_n) = Var\left[\frac{\summa (\gamma x_i + \varepsilon_i)x_i^3}{\summa x_i^4}\bigg|x_1, \dots, x_n\right] \underset{\text{i.i.d}}{=} \frac{\summa x_i^6 Var[\varepsilon_i|x_i]}{\left(\summa x_i^4\right)^2} 
	\end{align*}
	If $Var(\varepsilon|x) = const = \sigma^2$ then
	\begin{align*}
	Var(\tilde{\gamma}|x_1, \dots, x_2) = \sigma^2 \frac{\summa x_i^6}{\left(\summa x_i^4\right)^2} 
	\end{align*}
\end{enumerate}
\end{document}
